The Blessing of FedAvg's Linear Speedup in Federated Learning

### 2020-05-28 source/sec_intro.tex
Intro setup v0

% Our contributions
Our work aims to systematically fill the gap in the convergence analysis
of FedAvg and its accelerated variants. 
Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee all convex smooth objectives, which shows the convergence rate enjoys a linear speedup 
Furthermore, we studied a popular over-parameterized setting~\cite{liu2018accelerating} where the optimal solution can obtain zero training loss. In this scenario, we provide a novel accelerated FL algorithm improving
the convergence rate of FedAvg. Last but not least, we conduct extensive
evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.

>>>>

% Our contributions
Our work aims to systematically fill the gap in the convergence analysis
of FedAvg and its accelerated variants. 
Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee for all convex smooth objectives, which shows the convergence rate enjoys a linear speedup w.r.t.
the number of workers.
Furthermore, we studied a popular over-parameterized setting~\cite{liu2018accelerating} where the optimal solution can obtain zero training loss. In this scenario, we provide a novel accelerated FL algorithm improving
the convergence rate of FedAvg. Last but not least, we conduct extensive
evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.

### 2020-05-28 source/sec_problem.tex
Intro setup v0

\section{Problem Setup}

\subsection{The Federated Learning Problem}

The federated learning problem, given local objectives $F_k$ and device weights $p_k$, is to minimize

\begin{align}
	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\}
	\label{eq:problem}
\end{align}

where $N$ is the number of devices, and device weights $p_k$ satisfy $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the $k$-th device
holds the $n_k$ data points: $\xi_{k,1}, \xi_{k,2}, \dots, \xi_{k,n_k}$ . We assume that the local objective $F_k(\cdot)$ is defined by

\begin{align}
F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \xi_{k, j}\right)	
\label{eq:localloss}
\end{align}

where $\ell$ is some loss function. Thus, in the federated learning setting, each device stores its local data, which gives rise to its own local objective $F_k$. A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e. $F_k(w)$ may have different minimizers $w^{\ast}_k$. Moreover, each local device only has access to stochastic gradients based on data stored on the local device. In this paper, we investigate local objectives that are convex or strongly convex and smooth, although convergence results for non-smooth objectives are also given in the appendix.

\subsection{The FedAvg Algorithm}

Despite the heterogeneity in local objectives, the federated learning
problem can still be solved efficiently in a distributed manner with
the simple Federated Average (FedAvg) algorithm, which consists of
the iterations

\begin{align}
\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}g_{t,k}\\
\vw_{t+1}^{k} & =\begin{cases}
\vv_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\vv_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align}

>>>>

\section{Setup}

% \subsection{The Federated Learning Problem}

We study the federated learning problem
\begin{align}
	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\},
	\label{eq:problem}
\end{align}
where $N$ is the number of local devices (users/nodes/workers) and $p_k$ is the $k$-th device's weight satisfying $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. 
In $k$-th local device, there are a finite number of $n_k$ data points:
$\vx_k^1, \vx_k^2, \dots, \vx_k^{n_k}$.  
The local objective $F_k(\cdot)$ is defined as:
% \begin{align}
$F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \vx_{k, j}\right)$,
% \label{eq:localloss}
% \end{align}
where $\ell$ denotes a user-specified loss function. In federated learning, each device stores its local data, which gives rise to its own local objective $F_k$. Note that we do not make any assumption on the
data distributions of each local device. The local minimum
$\vw^{*}_k= \argmin_{\vw \in R^d} F_k(\vw)$ can be drifted far from
the global minimum of \eq{\ref{eq:problem}}, which quantifies the 
heterogeneity of the problem.
In this paper, we systematically investigate the convergence results, 
when the local objectives are convex smooth, strongly convex and convex but
non-smooth.

% A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e. $F_k(w)$ may have different minimizers $w^{\ast}_k$. Moreover, each local device only has access to stochastic gradients based on data stored on the local device. In this paper, we investigate local objectives that are convex or strongly convex and smooth, although convergence results for non-smooth objectives are also given in the appendix.

\subsection{The Federated Averaging (FedAvg) Algorithm}
Although the local objectives are heterogeneous, federated learning
problem can be solved efficiently in a distributed manner with
the Federated Average (FedAvg) algorithm, which consists of
the following procedures at $t$-th round: 
\begin{align}
\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}g_{t,k}\\
\vw_{t+1}^{k} & =\begin{cases}
\vv_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\vv_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align}
